---
title: "Worksheet: Data Wrangling Worksheet"
subtitle: "STT 2860 Intro to Data Mgmt & Visualization" 
author: "PUT YOUR NAME HERE"
date: "`r format(Sys.time(), '%A, %B %d, %Y @ %I:%M %p')`"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 4 
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

### Packages Used

```{r}
# load all packages used here
library(nycflights13)
library(ggplot2)
library(dplyr)

```

## Display data

In this worksheet, we will use the `flights` data frame in the `nycflights13` package. Use the `glimpse()` to display the `flights` data

```{r }

```




## `filter` rows 

- We begin by focusing only on flights from New York City to Portland, Oregon. 

    - The dest destination code (or airport code) for Portland, Oregon is "PDX". 
    - Run a code to ensure that only flights heading to Portland are chosen here:



```{r }

```





## `filter` rows 

- We filter flights for all rows that
    
    - departed from JFK and 
    - were heading to Burlington, Vermont ("BTV") or Seattle, Washington ("SEA") and
    - departed in the months of October, November, or December.
    


```{r }

```







## `filter` rows 


Lets filter rows corresponding to flights that didn’t go to Burlington, VT or Seattle, WA. 


```{r }

```






## `filter` rows 

We filter flights for all rows that were heading to the following ariports: `SEA`, `SFO`, `PDX`, `BTV` and `BDL`. 


```{r }

```





## `summarize` variables

Let’s calculate two summary statistics (mean and standard deviation) of the `temp` temperature variable in the `weather` data frame from `nycflights13` package. 


```{r }

```







## `group_by` rows

Say instead of a single summary statistics (mean and standard deviation) of temperature for the whole year, you would like 12 summary statistics, one for each of the 12 months separately. 
    

```{r }

```




## Grouping by more than one variable

Provide the frequency table by grouping by `origin` and `month`. Hint: use the `n()` function. 
 
 
```{r }

```







## `mutate` existing variables


For example, we can create a new variable by converting temperatures from $^\circ$F to $^\circ$C using the formula
$$\text{temp in C}=\frac{\text{temp in F}-32}{1.8}$$

## `mutate` existing variables

We can apply this formula above to the `temp` variable using the `mutate()` function from the `dplyr` package.


```{r }

```





## `mutate` existing variables

Let’s now compute monthly average temperatures in both $^\circ$F and $^\circ$C. 


```{r }

```





## `arrange` and sort rows


Suppose we are interested in determining the most frequent destination airports for all domestic flights departing from New York City in 2013. 



```{r }

```





## `select` variables

We’ve seen that the `flights` data frame in the `nycflights13` package contains 19 different variables. However, say you only need two of these 19 variables, say `carrier` and `flight`. You can `select()` these two variables:


```{r }

```


Let’s say instead you want to drop, or de-select, certain variables. For example, lets say we want to remove the `year` in the `flights` data frame. We can deselect year by using the `-` sign:


```{r }

```






***
```{r}
sessionInfo()
```